Post-Disturbance Tree Community Trajectories in a Neotropical Forest

### 2018-05-03 WholePlotTrajectories.tex
small corrections

In tropical forests communities are shaped by a constant range of
disturbance that changes the abiotic environment, as the light, heat and
water fluxes, and the biotic interaction and competitive pressure. The
cornerstone of tropical forests ecology is then to understand their
response to disturbance and its mechanisms
\citep{White2001, Chazdon2003a}. Forests response to disturbance has
been largely studied through structural parameters rapid and convenient
to measure, as aboveground biomass, tree height or stem density. These
structural parameters have then been sucessfully modeled and assessed
the maintenance of ecosystems processes and services
\citep{Denslow2000, Blanc2009, Rutishauser2016}. However the response of
tree species diversity although it is determinant of ecosystems
productivity, stability and functioning \citep{Tilman2014} and it would
be most probably impacted by the biotic and abiotic changes after
disturbance \citep{CazzollaGatti2014}.

In the short-term diversity dynamics demonstrated negligible or even
positive impacts of disturbance on communities diversity, which have
been formalized by the intermediate disturbance hypothesis (IDH) stating
a maximized species diversity at intermediate disturbance intensity
\citep{Molino2001, Kariuki2006a, Berry2008a}. However, validations of
the IDH in the long term remain scarce and mainly rely on the analysis
of species richness that gives limited or misleading information on
forests recovery and functioning \citep{Martin2015, Chaudhary2016}. More
releveant monitoring would encompass communities composition, that are
crucial for conservation issues, and evenness, that revealing ecological
rules underlying communities' structure
\citep{Magurran1988, Lavorel2002, Bellwood2006}. Furthermore, the
functional approach accounting for species biological attributes would
be insightful as it reveals species fitness and ecosystems functioning
\citep{Violle2007b, Moretti2009, Baraloto2012a, Scheiter2013}. In that
respect major functional traits related to species ecology and
performance were largely adopted through a relevant framework
\citep{Diaz2005, Villeger2008a}. The functonial trait-based approach,
for example, highlighted in tropical rainforests the environmental
filters fostering disturbance resistant species with rapid growth and
efficient resources acquisition \citep{Molino2001, Haddad2008}. This
translated by shifts from ``conservative'' slow-growing species dealing
with scarce resources, to ``acquisitive'' fast-growing species with
rapid and efficient use of abundant resources
\citep{TerSteege2001, Reich2014, Herault2011}. It was mirrored by shifts
in key functional traits related to resource acquisition (leaf area,
density and chlorophyll content, and stem specific gravity and bark
thickness), tree growth and reproduction life history traits (seed mass
and maximum height) \citep{Wright2004, Westoby2006a, Chave2009b}. A
proper monitoring of communities response should therefore encompass
taxonomic and functional diversity and composition measures to test the
validity of the IDH in the long term for tropical forests, and clarify
the resilience of communities evenness, composition,and also
functioning. The trajectories followed by all these facets would
highlight the role of deterministic processes, like competitive
exclusion or abiotic selection, and the communities' convergence
maintaining intrinsinc differences in diversity and composition,which is
as much insgights for future adaptive conservation strategies
\citep{Adler2007}.

@@

Paracou station in French Guiana (5°18'N and 52°53'W) is located in a
lowland tropical rain forest corresponding to a tropical wet climate
with mean annual precipitation averaging 2980 mm.y\textsuperscript{-1}
(30-y period) with a 3-months dry season (\textless{} 100
mm.months\textsuperscript{-1}) from mid-August to mid-November, and a
one-month dry season in March \citep{Wagner2011}. Elevation ranges
between 5 and 50 m and mean annual temperature is 26°C. Soils correspond
to thin acrisols over a layer of transformed saprolite with low
permeability generating lateral drainage during heavy rains. The
experiment corresponds to a network of twelve 6.25ha plots that have
undergone a gradient of three logging, thinning and fuelwood treatments
(Table \ref{tab:Tab1}). Disturbance treatments were attributed according
to a randomized plot design with three replicate blocks of four plots.
The disturbance corresponds to averages of 10 trees removed per hectare
with a diameter at 1.3 m height (DBH) above 50 cm for treatment 1 (T1),
32 trees/ha above 40 cm DBH for treatment 2 (T2) and 40 trees above 40
cm DBH for treatment 3 (T3). Treatments T2 and T3 besides included the
thinning of trees by poison girdling \citep{Blanc2009}.

@@

The study site corresponds to a tropical rainforest with a dominance of
Fabaceae, Chrysobalanaceae, Lecythidaceae and Sapotaceae botanical
families. In the twelve experimental plots of the experiment, all trees
above 10 cm DBH were mapped and measured annually since 1984. During
inventories, trees were first identified with a vernacular name assigned
by the field team, and afterward with a scientific name assigned by a
botanist during regular botanical campaigns. In 1984, specific
vernacular names were given to 62 commercial or common species whereas
other less common species were identified under two identifiers only
separating trees and palm trees. The botanical campaigns carried every 5
to 6 years to identify all trees at the species level only started in
2003 and identification practices varied among plots and successive
campaigns. This raised methodological issues as vernacular names usually
correspond to different botanical species, resulting in significant
taxonomic uncertainties that had to be propagated to composition and
diversity metrics throught a Bayesian framework. The uncertainty
propagation was done by the replenishment of inventories completed at
genus level from real incomplete ones on the basisof
vernacular/botanical names association.

Vernacular names were replaced through multinomial trials
\(M_v\Big(\big[s_1, s_2, …, s_N\big],\big[\alpha_1, \alpha_2,…, \alpha_3\big]\Big)\)
based on the observed association probability
\(\big[\alpha_1, \alpha_2,…, \alpha_3\big]\) between each vernacular
name \emph{v} and the species \(\big[s_1, s_2, …, s_N\big]\) recorded in
the inventory.

To avoid remaining identification caveats and consider complete
inventories, the simulated inventories were then reported at genus. See
appendix 1 and \citet{Aubry-Kientz2013} for the detailed methodology.

@@

The trait database came from the BRIDGE project {[}\^{}1{]} where traits
values were assessed from a selection of individuals located in nine
permanent plots in French Guiana, including two in Paracou. Missing
trait values were filled using multivariate imputation by chained
equation (mice) restricted to samples pertaining to the next higher
taxonomic level, in order to account for the phylogenetic signal of the
functional traits. The dataset comprised 294 botanical species
pertaining to 157 botanical genera.

@@

{[}\^{}1{]} http://www.ecofog.gf/Bridge/

@@

\bibliographystyle{mee}
\bibliography{references.bib}

>>>>

In tropical forests communities are shaped by a constant range of
disturbance that changes the abiotic environment, as the light, the heat
and water fluxes, and the biotic interaction and competitive pressure.
The cornerstone of tropical forests ecology is then to understand their
response to disturbance and its mechanisms
\citep{White2001, Chazdon2003a}. Forests response has been largely
studied through structural parameters rapid and convenient to measure,
as aboveground biomass, tree height or stem density. These structural
parameters have then been sucessfully modeled and allowed to assess the
maintenance of ecosystems processes and services
\citep{Denslow2000, Blanc2009, Rutishauser2016}. However the response of
tree species diversity, which is determinant of ecosystems productivity,
stability and functioning \citep{Tilman2014} and would be most probably
impacted by post-disturbance environmental changes
\citep{CazzollaGatti2014}, remains unclear.

In the short-term diversity dynamics demonstrated negligible or even
positive impacts of disturbance on communities diversity, which have
been formalized by the intermediate disturbance hypothesis (IDH) stating
a maximized species diversity at intermediate disturbance intensity
\citep{Molino2001, Kariuki2006a, Berry2008a}. However, validations of
the IDH in the long term remain scarce and mainly rely on the analysis
of species richness that gives limited or misleading information on
forests recovery and functioning \citep{Martin2015, Chaudhary2016}. More
releveant monitoring would encompass communities composition, that is
crucial for conservation issues, and evenness, that reveals ecological
rules underlying communities' structure
\citep{Magurran1988, Lavorel2002, Bellwood2006}. Furthermore, the
functional approach accounting for species biological attributes would
be insightful as it reveals species fitness and ecosystems functioning
\citep{Violle2007b, Moretti2009, Baraloto2012a, Scheiter2013}. In that
respect major functional traits related to species ecology and
performance were largely adopted through a relevant framework
\citep{Diaz2005, Villeger2008a}. The functional trait-based approach,
for example, highlighted in tropical rainforests the environmental
filters fostering disturbance resistant species with rapid growth and
efficient resources acquisition \citep{Molino2001, Haddad2008}. This
translated by shifts from ``conservative'' slow-growing species dealing
with scarce resources, to ``acquisitive'' fast-growing species with
rapid and efficient use of abundant resources
\citep{TerSteege2001, Reich2014, Herault2011}. It was mirrored by shifts
in key functional traits related to resource acquisition (leaf area,
density and chlorophyll content, and stem specific gravity and bark
thickness), tree growth and reproduction life history traits (seed mass
and maximum height) \citep{Wright2004, Westoby2006a, Chave2009b}. A
proper monitoring of communities response should therefore encompass
taxonomic and functional diversity and composition measures to test the
validity of the IDH in the long term for tropical forests, and clarify
the resilience of communities evenness, composition, and also
functioning. The trajectories followed by all these facets would
highlight the role of deterministic processes, like competitive
exclusion or abiotic selection, and the communities' convergence
maintaining intrinsinc differences in diversity and composition,which is
as much insgights for future adaptive conservation strategies
\citep{Adler2007}.

@@

Paracou station in French Guiana (5°18'N and 52°53'W) is located in a
lowland tropical rain forest corresponding to a tropical wet climate
with mean annual precipitation averaging 2980 mm.y\textsuperscript{-1}
(30-y period) with a 3-month dry season (\textless{} 100
mm.month\textsuperscript{-1}) from mid-August to mid-November, and a
one-month dry season in March \citep{Wagner2011}. Elevation ranges
between 5 and 50 m and mean annual temperature is 26°C. Soils correspond
to thin acrisols over a layer of transformed saprolite with low
permeability generating lateral drainage during heavy rains. The
experiment corresponds to a network of twelve 6.25ha plots that have
undergone a gradient of three logging, thinning and fuelwood treatments
(Table \ref{tab:Tab1}). Disturbance treatments were attributed according
to a randomized plot design with three replicate blocks of four plots.
The disturbance corresponds to averages of 10 trees removed per hectare
with a diameter at 1.3 m height (DBH) above 50 cm for treatment 1 (T1),
32 trees/ha above 40 cm DBH for treatment 2 (T2) and 40 trees above 40
cm DBH for treatment 3 (T3). Treatments T2 and T3 besides included the
thinning of trees by poison girdling \citep{Blanc2009}.

@@

The study site corresponds to a tropical rainforest with a dominance of
Fabaceae, Chrysobalanaceae, Lecythidaceae and Sapotaceae botanical
families. In the twelve experimental plots of the experiment, all trees
above 10 cm DBH were mapped and measured annually since 1984. During
inventories, trees were first identified with a vernacular name assigned
by the field team, and afterward with a scientific name assigned by a
botanist during regular botanical campaigns. In 1984, specific
vernacular names were given to 62 commercial or common species whereas
other less common species were identified under two identifiers only
separating trees and palm trees. The botanical campaigns carried every 5
to 6 years to identify all trees at the species level only started in
2003 and identification practices varied among plots and successive
campaigns. This raised methodological issues as vernacular names usually
correspond to different botanical species, resulting in significant
taxonomic uncertainties that had to be propagated to composition and
diversity metrics through a Bayesian framework. The uncertainty
propagation was done by the replenishment of inventories completed at
genus level from real incomplete ones on the basis of
vernacular/botanical names association.

Vernacular names were replaced through multinomial trials
\(M_v\Big(\big[s_1, s_2, …, s_N\big],\big[\alpha_1, \alpha_2,…, \alpha_3\big]\Big)\)
based on the observed association probability
\(\big[\alpha_1, \alpha_2,…, \alpha_3\big]\) between each vernacular
name \emph{v} and the species \(\big[s_1, s_2, …, s_N\big]\) recorded in
the inventory. See appendix 1 and \citet{Aubry-Kientz2013} for the
detailed methodology.

@@

The trait database came from the BRIDGE project \footnote{http://www.ecofog.gf/Bridge/}
where trait values were assessed from a selection of individuals located
in nine permanent plots in French Guiana, including two in Paracou.
Missing trait values were filled using multivariate imputation by
chained equation (mice) restricted to samples pertaining to the next
higher taxonomic level, in order to account for the phylogenetic signal
of the functional traits. The dataset comprised 294 botanical species
pertaining to 157 botanical genera.

@@

@@

\bibliographystyle{mee}
\makeatletter
% The filename has .bib extension the must be eliminated
\filename@parse{references.bib}
% parse stores the file name in base. Extension starts at the first dot, so don't use dots in file names.
\bibliography{\filename@base}
\makeatother
